### 2021-01-06 draft/sections/visual.tex
q

\documentclass[../main.tex]{subfiles}
\begin{document}

\end{document}