### 2021-03-14 paper/scraps.tex
base space

\begin{figure}[tb]
    \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
    \includegraphics[width=\columnwidth]{monoid_maps.png}
    \caption{Monoids support partial ordering, such as the two emojis with tilted mouths being at the same rank above and below emojis with different mouth positions. The monotonic map on the right is a meta ordering of the partial order.  Figure inspired by definition 1.59 diagram in Spivak and Fong's An Invitation to Applied Category Theory \cite{fongInvitationAppliedCategory2019}}
    \label{fig:data_partial_order}
  \end{figure}